Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-01 ms.tex
working on discussion despite myself

Figure \ref{fig:bh_growth} shows that across all black hole masses, Compton heating alone cannot prevent thermal instabilities ($\dot{M}_{\rm C} > \dot{M}_{\rm TI}$), as will described more below.  In the case of an old stellar population which produces insufficient heating to support steady accretion, this large scatter in Fig \ref{fig:Miller} can in principle also be understood by the wide range of gas masses experienced over the course of a cyclic episode between instabilities and periodic nuclear outbursts (e.g.~\citealt{Ciotti+10}).  The cyclic model also predicts that high $L_{X}$ nuclei should be more common in galaxies with younger stellar ages.  This appears in tension with the finding by \citet{Pellegrini10} that galaxies with younger stellar populations are observed to host less X-ray luminous nuclei, at least for stellar ages $<3$ Gyr.  This observation can also be understood in the stable accretion picture by the fact that young stars provide sufficient stellar wind heating to produce a thermally stable accretion state.

In general we find that core galaxies are more likely to possess higher accretion rates and potentially be thermally unstable.  This may be consistent with past studies which have found that core galaxies show a wide range of nuclear radio and X-ray luminosities, while cusp galaxies are confined below a threshold (e.g., \citealt{Bender+89, Pellegrini99, Capetti&Balmaverde05}).

GALEX observations have shown that a small fraction of early-type galaxies in the local universe have undergone recent ($< 1-2$ Gyr old) star formation episodes (\citealt{Donas+07}).

\subsection{Demographics of SMBHs}

>>>>

Figure \ref{fig:bh_growth} (top panel) shows that across all black hole masses, Compton heating alone cannot prevent thermal instabilities from developing in a steady-state accretion flow ($\dot{M}_{\rm C} > \dot{M}_{\rm TI}$).  Thus, if the stellar population is sufficiently old (or the star formation sufficiently intermittent) then nuclear gas could undergo cyclic oscillations of thermal instability and high accretion rates, followed by quiescent periods once gas has been consumed and/or AGN feedback becomes effective.  Such cycles are proposed to occur on large scales and long timescales in elliptical galaxies (\citealt{Ciotti+10}), an idea supported, e.g. , by GALEX observations showing that a small fraction of early-type galaxies in the local universe have undergone recent ($< 1-2$ Gyr old) star formation episodes (\citealt{Donas+07}).  If such cyclic behavior also occurs on small, parsec scales (e.g.) then the large scatter in the accretion rates and hence X-ray luminosities shown in Fig \ref{fig:Miller} could also be understood by the wide range of gas masses experienced over the course of a cyclic episode between instabilities and periodic nuclear outbursts (e.g.~\citealt{Ciotti+10}).  In support of this possibility, we find that core galaxies are more likely to possess higher accretion rates and potentially be thermally unstable.  Past studies have found that core galaxies show a wide range of nuclear radio and X-ray luminosities (suggesting cyclic accretion episodes), while cusp galaxies are confined below a threshold (e.g., \citealt{Bender+89, Pellegrini99, Capetti&Balmaverde05}), consistent with thermally-stable picture.

The cyclic model predicts that high $L_{X}$ nuclei should be more common in galaxies with younger stellar ages, a result which appears in tension with the finding by \citet{Pellegrini10} that galaxies with younger stellar populations are observed to host less X-ray luminous nuclei, at least for stellar ages $<3$ Gyr.  An alternative explanation for this observation within a thermally stable accretion picture is that young stars provide higher stellar wind heating, reducing the mass accretion rate and hence nuclear X-ray luminosity.

\subsection{Local Growth of SMBHs }
